---
title: "R in 3 Months: Week 7 (Tidy Data)"
output:
  xaringan::moon_reader:
    css: ["style.css", "default"]
    lib_dir: libs
    df_print: paged
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE, eval = TRUE}
options(htmltools.preserve.raw = FALSE)
library("tidyverse")
library("wakefield")
library(palmerpenguins)
library(janitor)
```



# Agenda

--

1. Housekeeping

--

1. Tidy Data

--

1. Exercise

--

1. More Complicated Tidying Examples

--

1. Next Week

---
class: inverse, center, middle

# Housekeeping

???

Discuss week

---

class: inverse, center, middle

# Tidy Data

---

class: inverse, center, middle

# What questions do you have about tidy data?

---

## What is tidy data?

> "Tidy datasets are all alike, but every messy dataset is messy in its own way"
> 
> Hadley Wickham

--

Tidy data is both a philosophical approach to wrangling data - and a requirement for using parts of the tidyverse.

---

## Defining tidy data

From the [tidy data vignette](https://tidyr.tidyverse.org/articles/tidy-data.html#defining):

--

1. Every column is a variable.

--

1. Every row is an observation.

--

1. Every cell is a single value.



---

# Tidy data and `ggplot2`

`ggplot` needs data in a tidy format. This dataset is the exact opposite of what {ggplot2} needs for a bar chart.

```{r, echo=FALSE}
msleep %>% 
  drop_na(vore) %>% 
  count(vore) %>% 
  pivot_wider(names_from = vore,
              values_from = n)
```

---

## Tidy data and `ggplot2`

The previous dataset is wide. What `ggplot2` needs is long.

```{r, echo=FALSE}
msleep %>% 
  drop_na(vore) %>% 
  count(vore)
```

---

## Is this dataset wide or long?

`data_mike_week_3`

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
data_mike_week_3 <- read_csv(here::here("data", "mike-data-week-3.csv")) %>% 
  slice(1:10) 
data_mike_week_3
```

---

## Is this dataset wide or long?

`wakefield_dataset`

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(wakefield)

r_data_frame(
  100,
  name = name(),
  age = age(),
  gender = gender_inclusive(prob = c(0.5052, 0.4808, 0.017)),
  religion = religion()
) %>% 
  mutate(across(where(is.factor), ~as.character(.x)))
```

---

## Is this dataset wide or long?

`tidycensus`

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(tidycensus)
library(tidyverse)

acs_poverty <- get_acs(year = 2019,
        geography = "county",
        state = "OR",
        variables = c("n_in_poverty_male" = "S1701_C02_011",
                      "n_in_poverty_female" = "S1701_C02_012",
                      "n_in_poverty_under_18" = "S1701_C02_002"))

acs_poverty
```

---


## Is this dataset wide or long?

`tidycensus v2`

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
acs_poverty_wide <- get_acs(year = 2019,
                            geography = "county",
                            state = "OR",
                            variables = c("n_in_poverty_male" = "S1701_C02_011",
                                          "n_in_poverty_female" = "S1701_C02_012",
                                          "n_in_poverty_under_18" = "S1701_C02_002"),
                            output = "wide")

acs_poverty_wide
```

---

## Is this dataset wide or long?

`wikipedia_calendars`

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
read_csv(here::here("data", "wikipedia-tables.csv"))
```

---

# `pivot_longer()`

```{r eval = FALSE}
your_data %>% 
  pivot_longer(cols = columns_to_pivot,
               names_to = "name_column_name",
               values_to = "value_column_name")
```



---

## `pivot_longer()`

```{r}
relig_income
```

---

## `pivot_longer()`

```{r}
relig_income %>% 
  pivot_longer(cols = -religion,
               names_to = "income_group",
               values_to = "n")
```



---

class: inverse, center, middle

# Exercise

---

## Exercise

Copy the code from [this gist](https://gist.githubusercontent.com/dgkeyes/1d5ceb8b607266f1d197a218dd5c6882/raw/3591c9dc6e02ee78e3471bc40753cc948456c4d1/tidy-data-exercise.R) into a new R script file. Then tidy the `gapminder_wide` data.


---

class: inverse, center, middle

# More complicated tidying

---

### pivot_longer(names_sep)

It's possible to tell `pivot_longer()` to split column names into two "name" columns.

```{r echo=FALSE, message=FALSE, warning=FALSE}
gapminder_wide <- read_csv("https://raw.githubusercontent.com/swcarpentry/r-novice-gapminder/gh-pages/_episodes_rmd/data/gapminder_wide.csv")

gapminder_wide
```


---

## Why is wide data sometimes desirable?

Do you have any ideas?

--

- If you're making beautiful tables, you likely want wide data.

--

- You can't mix different data types (eg age and religion)

--

- Joining datsets with `left_join()` sometimes requires wide data


---

# Next Week: Separating columns and multiple choice responses

Next week we will talk about survey datasets and separating columns that contain multiple choice (i.e. check all that apply) items.

---

## Next Week

--

There are two big topics in the next week:

--

1. Reducing repetitive code with user-defined functions

--

1. Joining together datasets by shared columns (i.e. data merging)

--

These skills will equip you to deal with more complex and larger scale data wrangling

---

## Project assignment 

Use the `pivot_longer()` function to tidy your data and visualise it with {ggplot2}

<!-- Example from the video https://gist.github.com/charliejhadley/80e3182e76639022b08b75dfa426ff84 -->
