---
title: "R in 3 Months<br>Week 8<br>More Data Tidying"
output:
  xaringan::moon_reader:
    css: ["style.css", "default"]
    lib_dir: libs
    df_print: paged
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE, eval = TRUE}
options(htmltools.preserve.raw = FALSE)
library(tidyverse)
library(readxl)
library(janitor)
library(fivethirtyeight)
library(here)
library(gapminder)
library(patchwork)
library(tweetrmd)
```


<!---- PLAN

Other types of tidy data - surveys! separate()

- Compare SurveyMonkey & Qualtric which make multiple columns, vs Google Form uses one column
--- Make a thread comparing 


--->

# Agenda

--

1. Housekeeping

--

1. One tip

--

1. Other functions for tidying data

--

1. David and Charlie tidy survey data

--

1. Next week

---

# Housekeeping


---

class: inverse, center, middle

# One tip

---

## pivot_longer() complains it can't combine

Often you'll see this error:

```{r}
# Error in `pivot_longer_spec()`:
#
# ! Can't combine `ACO1` <double> and `ACO13` <character>.
# Run `rlang::last_error()` to see where the error occurred.
```

Let's walk through an example

---

## pivot_longer() complains it can't combine

We want to combine together all of the columns from `ACO1:ACO28`

```{r message=FALSE, warning=FALSE, echo=FALSE}
data_performance <- read_csv(here::here("data", "Performance_Year_Financial_and_Quality_Results_2020.csv")) %>% 
  select(contains(c("ACO"))) %>% 
  mutate(across(ACO1:ACO28,
                ~  ifelse(.x %in% c("^", "~", "-"), 0, .x))) 
data_performance
```


---
class: inverse, center, middle

# Other functions for tidying data

---

## separate()

The `separate()` function is useful when a column contains **multiple variables**.

```{r, eval=TRUE}
location_data <- tribble(
  ~id, ~address,
  1, "Las Vegas, USA",
  2, "Bristol, UK",
  3, "Kassala, Sudan"
)
location_data
```


---

## separate()

```{r, eval=TRUE}
location_data %>% 
  separate(col = address,
           into = c("city",
                    "country"),
           sep = ", ")
```

---

## separate_rows()

The `separate_rows()` function is useful when a column contains **multiple observations of the same variable**.

---

## separate_rows()

Which of these devices do you own? Please select all that apply.

--

```{r, eval=TRUE, echo=TRUE}
device_ownership <- tribble(
  ~name, ~devices_owned,
  "Charlie", "Smart TV, Cell phone",
  "Mohammad", "Cell phone",
  "Christina", "Smart TV, Games Console, Cell phone"
)
device_ownership
```

---

## separate_rows()

```{r, eval=TRUE, echo=TRUE}
device_ownership %>% 
  separate_rows(devices_owned,
                sep = ", ")
```

---

## Tidy data definition

--

1. Every column is a variable.

--

1. Every row is an observation.

--

1. Every cell is a single value.


---

class: inverse, center, middle

# Multiple choice survey questions

---

class: middle

```{r echo = FALSE}
tweet_embed("https://twitter.com/charliejhadley/status/1519696924328214528",
            align = "center")
```

---

# Let's tidy the survey data!

David and I will take turns to read in and tidy the survey data from:

- Survey Monkey

- Qualtrics

- Google Forms

---

class: inverse, center, middle

# User-defined functions

---

# User-defined functions

During this week's videos you learned about creating user-defined functions.

This is a really powerful part of learning a programming language like R. They allow you to re-use code to do multiple things, including:

- Clean data files that share the same format (eg yearly reports)

- Create data visualisations of different subsets of your data, eg a line chart showing visitor numbers for each exhibit

- Generate the same report - including lots of different kinds of charts and tables - different subsets of data. This is something we do in consulting projects, eg for [ORS Impact Reports](https://funderscommittee.org/2020censusstatereports/).

---

class: inverse, center, middle

# Next Week

---

# Next week

- Advanced data visualisation with {ggplot2}, including using {ggtext} for direct labelling

- No new project assignment

- The week of Monday 21st November is another catch-up week