---
title: "R in 3 Months: Week 6<br>(Git + GitHub) HELLO!"
output:
  xaringan::moon_reader:
    css: ["style.css", "default"]
    lib_dir: libs
    df_print: paged
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, eval = TRUE}
options(htmltools.preserve.raw = FALSE)
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)
library("tidyverse")
library("wakefield")
library(palmerpenguins)
library(janitor)
library(ggtext)
```


<!---- PLAN

THESE SLIDES USE sli.do FOR POLLS!

USE THIS LINK: https://app.sli.do/event/ndcJPnZgvf18CgCbCp93hn/live/polls

--->

```{r}

```

# Agenda

--

1. Housekeeping

--

1. Four Tips

--

1. GitHub

--

1. Next Week

---

# Housekeeping

--

- Catch-up Week

--

---

class: inverse, center, middle

# Four Tips

---

class: inverse, center, middle

# Data Viz Examples

---

### (POLL) How would we change the colors in the chart?

```{r, eval=FALSE}
penguins %>% 
  drop_na(flipper_length_mm, bill_length_mm) %>% 
  ggplot(aes(flipper_length_mm,
             bill_length_mm)) +
  geom_point(aes(color = ifelse(bill_length_mm > 40, "red", "black")))
```

.pull-left[
```{r, eval=TRUE, echo=FALSE, out.height='350px'}
penguins %>% 
  drop_na(flipper_length_mm, bill_length_mm) %>% 
  ggplot(aes(flipper_length_mm,
             bill_length_mm)) +
  geom_point(aes(color = ifelse(bill_length_mm > 40, "red", "black")))
```
]

.pull-right[
<center'><a href='https://app.sli.do/event/ndcJPnZgvf18CgCbCp93hn/live/polls'>app.sli.do/event/ndcJPnZgvf18CgCbCp93hn/live/polls</a></center>
]


---

class: inverse, center, middle

# Git + GitHub

???

- Have to save files first

---

# GitHub Desktop

The git tab in RStudio is great, it will suit most of your needs.

GitHub Desktop provides some additional tools and capabilities that might be useful for more advanced users.

--

If you can't get RStudio and git to work together then consider trying GitHub Desktop instead.

---

# Editing files in the browser on GitHub

The GitHub website gives you the ability to edit files directly.

Technically, this is exactly the same as making a commit.

Remember you will need to **pull** the changes onto your local machine.

---

class: inverse, center, middle

# Group Work

---

## Group Work

In your groups we'd like you to discuss your datasets with one another and to take a look at your charts.

- Please take it in turns to share your screen

- Share something you're struggling (or want) to do with your data. 

    - Ask if anyone else in the team needs to do something similar
    
    - Discuss potential solutions
    
    - Add your struggles (and possible solutions) [to this poll](https://app.sli.do/event/ndcJPnZgvf18CgCbCp93hn/embed/polls/4cefc4be-4c0d-42af-a6e9-ace6290f3ab1)

--

- Show a chart you've built to your group 

    - Describe something you'd like to add to your chart
    
    - Discuss potential solutions
    
    - Add your struggles (and possible solutions) [to this poll](https://app.sli.do/event/ndcJPnZgvf18CgCbCp93hn/embed/polls/2c788ef4-24df-499c-b614-7c6606d79874)
 
???

This is what we did last year:

We need to figure this out. Last time we did:

- In a group, nominate one person to create a new project, add a local Git repository, and then push it to GitHub.

- Have the nominated person share their screen and then walk through the steps as a group
  
- Add any questions that come up to [this discussion thread](https://rfortherestofus.com/groups/r-in-3-months-fall-2021/forum/discussion/general-git-github-questions/).

---

class: inverse, center, middle

## Git + GitHub Questions

???

Need to save first in order to commit

Can only commit items that have been modified since last commit


---

# Next Week
<!-- Charlie's example video was built with this gist https://gist.github.com/charliejhadley/b2bc9af0853779093f1ce7308687ef48 -->

--

1. Advanced data wrangling

--

1. The main thing to make sense of is the concept of tidy data (it's a hard concept, but worth the effort to figure it out!)

--

1. Project assignment: figure out ways your data is not tidy and tidy it (sounds simple, but it's not!).

--

1. You'll work in your GitHub repository from now on so you'll submit a link to the code in your repo for Charlie's review





